---
title: "Task1"
author: ' Arijit Guchhait'
date: "2025-01-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The main purpose of this task is to check that you can get it to load and run and knit it to html.
Once that's working, have a go at the embedded exercises. Exercise 7 is just bonus.
At the end, knit it to html and upload both files (Rmd and html) on blackboard (not zipped).

## R and the RStudio IDE

R is a free software environment for statistical computing and graphics. We will use the RStudio IDE (Integrated Development Environment). This has a number of nice features which we'll cover as we go through - identify the following elements:

  * Source Window
  * Console
  * Help
  * Environment

## Exercise 1a                
Create variables for the following information:
name (character): 
age (numeric): Your age
is_local (logical): Do you live in Leicester? (TRUE/FALSE)

```{r Exercise 1a}
name <- 'Arijit Guchhait'
age <- 32
is_local <- TRUE
```

## Exercise 1b
Is it possible to create a vector containing the values for your three variables? If not, create a named list containing the values for your three variables and print it.
```{r Exercise 1b}
#Vector only accept same datatypes
x <- list(a=name,b=age,c=is_local)
x
```


## Exercise 2

Create a data frame called "students" with the following columns:
Name (character): Names of students (include your name and at least three more names)
Age (numeric): Ages of students
Grade (factor): Grades of students (with levels: A, B, C)
Is_Local (logical): Whether each student lives in Leicester (TRUE/FALSE)


```{r Exercise 2}
students <- data.frame(
  Name=c('Arijit','Poulami','Rajita','Madhumita','Krittika'),
  Age= c(32,34,28,33,32),
  Grade = factor(c('A','A','B','B','C'),levels=c('A','B','C')),
  Is_Local =c(TRUE,FALSE,FALSE,FALSE,FALSE)
)
print(students)

```

## Exercise 3
Let R show how many times the different levels appear in the grade-column
```{r Exercise 3}
summary(students$Grade)
```


## Exercise 4

Extract and print the Age vector from the students data frame.
```{r Exercise 4}
print(students$Age)
```

## Exercise 5
Look up the function "mean()" with the help function.
Calculate the average age of the students from your dataframe and print it out. Use the pipe operator to do this calculation in one line of code.
```{r Exercise 5}
students$Age |> mean()
```

## Exercise 6
Why does the following code not run? Fix the Code (in the code chunk below named r "Exercise 6, solution"). Note, before you run the code, you need to get rid of the "#"s. You can do this by selecting everything and using the shortcut 'Ctrl Shift C'." (Commenting it out again works the same way).
```{r Exercise 6 the code with mistakes}
df <- data.frame(Name = c("John", "Jane", "Doe"), Age = c(25, 30,99))
print(df)

print("The problem was Age was not same dimension as name" )
age_in_10_years <- df$Age +10
mean(age_in_10_years)

```

```{r Exercise 6, solution}
df <- data.frame(Name = c("John", "Jane", "Doe"), Age = c(25, 30,99))
print(df)

print("The problem was Age was not same dimension as name" )
age_in_10_years <- df$Age +10
mean(age_in_10_years)
```



## Exercise 7* (bonus)
R contains some built-in data sets (that means they are included with the R installation and can be accessed without the need to load any external libraries.) One of them is "mtcars".
Print it out and explore it using the functions head(), View(), summary(), glimpse(), str().
Try to understand what the functions tell you about the data set.
Moreover, why should you not include the function "View()" when you try to knit your file?
```{r Exercise 7}
head(mtcars)
View(mtcars)
summary(mtcars)
dplyr::glimpse(mtcars)
str(mtcars)
#Because it will open in another windows
```


